---
title: "assignment7"
output: html_document
date: "2024-10-29"
---

How to do it?:

Open the Rmarkdown file of this assignment (link) in Rstudio.

Right under each question, insert a code chunk (you can use the hotkey Ctrl + Alt + I to add a code chunk) and code the solution for the question.

Knit the rmarkdown file (hotkey: Ctrl + Alt + K) to export an html.

Publish the html file to your Githiub Page.

Submission: Submit the link on Github of the assignment to Canvas

```{r}
library(tidyverse)
library(ggplot2)
df <- read_csv('Data Science Jobs Salaries.csv')
```

Make a plot using facet_wrap by one categorical variable (facet_wrap(~variable1))
```{r}
ggplot(df, aes(x = work_year, y = salary)) +
  geom_point() +
  facet_wrap(~ experience_level)
```

These maps compare the work year to salary and experiences of the employees.

Make a plot using facet_wrap by two categorical variables (facet_wrap(~variable1+variable2))
```{r}
ggplot(df, aes(x = work_year, y = salary)) +
  geom_point() +
  facet_wrap(~ experience_level + employment_type)
```
These maps compare the work year to salary and experiences of the employees as well as their employment type.

3. Make a plot using position ='dodge'.

```{r}
df %>% ggplot()+ 
  geom_bar(mapping=aes(x = work_year , fill = remote_ratio),
           position = 'dodge')
```
These bar graphs compare the work years to how many people were working remotely.

Make a plot using position ='fill'.

```{r}
df %>% ggplot()+ 
  geom_bar(mapping=aes(x = work_year , fill = remote_ratio), 
           position = 'fill')
```
These bar graphs compare the work years to how many people were working remotely.

5. Make a plot using geom_col.

```{r}
df %>% 
  ggplot()+ 
  geom_col(mapping=aes(x = experience_level, y = salary))
```
These bar graphs compare the experience level and the salary that the employment receives.

6. Change the theme of one of your plots.
```{r}
df %>% ggplot()+ 
  geom_bar(mapping=aes(x = work_year , fill = remote_ratio))+
theme_bw()
```

7. Make a plot using combo filter + ggplot
```{r}
df %>%
  filter(work_year == 2020, remote_ratio == 0)
```

```{r}
  ggplot(df, aes(x = work_year, y = remote_ratio)) +
  geom_col()
```

These two graphs show how many people worked remotely during each work year.

8. Make a plot using combo group_by + summarise + geom_col
```{r}
df %>% 
  group_by(experience_level) %>%
  summarise(averageSalary = mean(salary, na.rm=TRUE)) %>% 
  ggplot()+ 
  geom_col(mapping=aes(x = experience_level, y = averageSalary))
```

These graphs show the average salary for each level of experience.

9. Save a plot and put a link of the plot to your Github Webpage. To link a plot abc.png you just need to insert the following on the README.md

```


10. In this question, we will use the Covid19 vaccination data for the US. The data and data dictionary can be found at the below link:
```{r}
library(tidyverse)
df <- read_csv('vaccinations-by-manufacturer.csv')
```
Geom_point makes scatter plots, which plot two numeric (or Date) variables. A third categorical variable is often brought in for color. For example, we can compare different vaccine types used in the US by

```{r}
df %>% 
  filter(location=='United States') %>% 
  ggplot()+
  geom_point(mapping = aes(x = date, 
                                    y = total_vaccinations, 
                                    color = vaccine))
```

Do the follows.

Use geom_point to compare the vaccines used by date in Argentina in 2022.

```{r}
df %>% 
  group_by(date == 2022) %>% 
  filter(location == 'Argentina') %>% 
  ggplot()+
  geom_point(mapping = aes(x = date, y = total_vaccinations, color = vaccine))
```
Change geom_point to geom_line in the above codes to have the corresponding line plots.

```{r}
df %>% 
  group_by(date == 2022) %>% 
  filter(location == 'Argentina') %>% 
  ggplot()+
  geom_line(mapping = aes(x = date, y = total_vaccinations, color = vaccine))
```
11. (Using the same dataset as in 10.) Sometime we need to do some calculations on the data before plotting. For example, the below codes is to compare the total vaccines (including all types of vaccines) by date between the US and the EU

```{r}
d1 <- df %>% 
  filter(location %in% c('United States', 'European Union')) %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE))

# Plot
d1 %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, y=total_vac2, color=location))
```


```{r}
df %>% 
  filter(location %in% c('United States', 'European Union')) %>% 
  group_by(date, location) %>% 
  summarise(total_vac2=sum(total_vaccinations, na.rm=TRUE)) %>% 
  ggplot()+
  geom_point(mapping=aes(x=date, 
                         y=total_vac2, 
                         color=location))
```

Do the follows.

Use geom_point compare the total vaccines (including all types of vaccines) by date between countries/regions in Asia

```{r}
df %>% 
  filter(location %in% c('Hong Kong', 'Japan', 'South Korea', 'Nepal')) %>% 
  group_by(date, location) %>% 
  summarize(total_vac = sum(total_vaccinations, na.rm=TRUE)) %>% 
  ggplot()+
  geom_point(mapping=aes(x = date, y = total_vac, color = location))
```


Use geom_point compare the total vaccines (including all types of vaccines) by date between countries/regions in South America in 2021
```{r}
df %>% 
  filter(location %in% c('Argentina', 'Chile', 'Ecuador', 'Peru', 'Uruguay')) %>% 
  group_by(date = 2021, location) %>% 
  summarize(total_vac = sum(total_vaccinations, na.rm=TRUE)) %>% 
  ggplot()+
  geom_point(mapping=aes(x = date, y = total_vac, color = location))
```




