
---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 3.  Working with Data Frame. Base R Style"
---

***Note*:** *This assignment practices working with Data Frame using Base R.*

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment3.Rmd)) in Rstudio. 

- Right under each question, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas under Assignment 3. 

-------

Problems
-------
 \

1. Create the following data frame

| Rank 	| Age 	| Name  	|
|:------	|:-----	|:-------	|
| 0    	| 28  	| Tom   	|
| 1    	| 34  	| Jack  	|
| 2    	| 29  	| Steve 	|
| 3    	| 42  	| Ricky 	|

```{r}
df = data.frame(Rank = c('0','1','2','3'),
                Age = c(28,34,29,42),
                Name = c('Tom','Jack','Steve','Ricky'))
```


2. Use `read.csv` to import the Covid19 Vaccination data from WHO: [link](https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv).

2. Show the names of the variables in the data

```{r}
df = read.csv('https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv')
names(df)
```
3. How many columns and rows the data have?

```{r}
head(df)
```



4. How many missing values are there?  Show the missing values by columns.  What variable has the most number of missing values?
```{r}
sum(is.na(df))
```
```{r}
colSums(is.na(df))
```
```{r}
colMeans(is.na(df))
```




5. What is the class of the `date` column. Change the `date` columns to `date` type using the `as.Date` function. Show the new class of the `date` column. 
```{r}
class(df$date)
```
```{r}
df$date = as.Date(df$date)
```



6. Capitalize the names of all the variables
```{r}
names(df)[1] <- 'Date'
names(df)[2] <- 'State'
names(df)[3] <- 'Fips'
names(df)[4] <- 'Cases'
names(df)[5] <- 'Deaths'
```


7. Find the average number of cases per day. Find the maximum cases a day.  
```{r}
mean(df$Cases, na.rm=TRUE)
```
```{r}
max(df$Cases, na.rm=TRUE)
```


8. How many states are there in the data?
```{r}
length(table(df$State))
```


9. Create a new variable `weekdays` to store the weekday for each rows. 

```{r}
df$weekdays <- weekdays(df$Date)
```


10. Create the categorical variable `death2` variable taking the values as follows

- `has_death` if there is a death that day
-  `no_death` if there is no death that day

Find the frequency and relative frequency of `no_death` and `has_death`. 

```{r}
df$death2 <- ifelse(df$Deaths==0,
                    'no_death','has_death')
table(df$death2)
```


11. Find the first quartile (Q1), second quartile (Q2) and and third quartile (Q3) of the variable `death`. (Hint:  Use the `summary` function)

```{r}
summary(df$Deaths)
```


12. Create the categorical variable `death3` variable taking the values as follows

- `low_death` if the number of deaths smaller than the 25 percentile (Q1)

- `mid_death` if the number of deaths from Q1 to Q3

- `high_death` if the number of deaths greater than Q3

```{r}
library(dplyr)
```
```{r}
df$deaths3 <- case_when(df$Deaths<1171 ~ 'low_death',
                             df$Deaths<14912 ~ 'mid_death',
                             TRUE~'high_death')
table(df$deaths3)
```


13. Find the average cases in Rhode Island in 2021

```{r}
df <- df[df$Date_report>='2021-01-01',]
by(df1$New_cases, df1$WHO_region, mean)
```


14. Find the median cases by weekdays in Rhode Island in 2021 

```{r}
df$weekdays <- weekdays(df$Date)
df1 = df[df$state== 'Rhode Island',]
by(df1$deaths,df1$weekdays, median)
```


15. Compare the median cases in Rhode Island in June, July, August and September in 2021. 

```{r}
df$month = months(df$Date)
```

```{r}
df1 = df[df$month %in% c('June' , 'July' , 'August', 'September') & (df$date >= '01-01-2021') & (df$date <= '12-31-2021') ,]
by(df1$Cases, df1$month, median)
```


16.  Find your own dataset, import it and implement the following functions on the data

```{r}
df2 = read.csv("combined.csv")
```


- head
```{r}
head(df2)
```

- str
```{r}
str(df2)
```

- names
```{r}
names(df2)
```

- mean, min, max
```{r}
mean(df2$FlightNum, na.rm=TRUE)
```
```{r}
min(df2$FlightNum, na.rm=TRUE)
```
```{r}
max(df2$FlightNum, na.rm=TRUE)
```

- table
```{r}
table(df2$Month)
```

- is.na
```{r}
sum(is.na(df))
```

- colSums
```{r}
colSums(is.na(df2))
```

- class
```{r}
class(df2$FlightNum)
```

- cor
```{r}
cor(df2$Month, df2$FlightNum)
```

- by
```{r}
by(df2$FlightNum, df2$FlightNum, FUN = mean)
```


