---
title: "Midterm"
output: ioslides_presentation
date: "2024-11-12"
---


```{r}
library(gganimate)
library(ggplot2)
library(tidyverse)
```

I. Data Wranggling
Download the data file hdd0318cy.sas7bdat.

Use read_sas in library haven to read the data.
```{r}
library(haven)
df <-read_sas('hdd0318cy.sas7bdat')
raw_data=df
n=nrow(df)
print(raw_data)
```

```{r}
# find the number of rows of the data
n = nrow(df)

# subset 1000 rows of the data
df1 = df[sample(1:n, 1000), ]
```


Filter the data to have only patients of the year 2018 (yod=18)
```{r}
df1 = raw_data %>% 
  filter(yod==18)
```


Select to work with only following variables:

                      "yod", "payfix","pay_ub92","age",  
                      "sex","raceethn","provider","moa", 
                      "yoa","mod","admtype", "asource" , 
                      "preopday" ,"los", "service" , "icu","ccu",    
                      "dispub92", "payer"  ,"drg","trandb", 
                      "randbg","randbs","orr", "anes","seq",   
                      "lab","dtest", "ther","blood","phar", 
                      "other","patcon","bwght","total","tot" ,  
                      "ecodub92","b_wt","pt_state","diag_adm","ancilar" ,
                      "campus","er_fee","er_chrg","er_mode","obs_chrg",
                      "obs_hour","psycchrg","nicu_day"
Notice: You may want to save the current data to your computer for easy access later. To save the data file use write_csv(df, 'midterm.csv'), for example. Also notice that, empty values in the data before writing to csv may turn to NAs later when you re-read the file.
```{r}
write_csv(df1, 'midterm.csv')
```


```{r}
df1 <- df1 %>%
  mutate(sex = case_when(
    sex == 1 ~ "Male",
    sex == 2 ~ "Female"
  ))
df1 <- df1 %>% 
  mutate(provider = as.numeric(provider),
         provider = case_when(
    provider == 7201 ~ 'Newport',
    provider == 7202 ~ 'St. Joseph Health Services of RI',
    provider == 7203 ~ 'Memorial',
    provider == 7204 ~ 'Miriam',
    provider == 7205 ~ 'Rhode Island Hospital',
    provider == 7206 ~ 'Roger Williams',
    provider == 7209 ~ 'South County',
    provider == 7210 ~ 'Kent County',
    provider == 7211 ~ 'Westerly',
    provider == 7212 ~ 'Rehab of RI',
    provider == 7213 ~ 'Landmark Medical Center',
    provider == 7214 ~ 'Women and Infants',
    provider == 7215 ~ 'Bradley',
    provider == 7216 ~ 'Butler',
  ))
df1$total <- as.numeric(df1$total)
df1$icu <- as.numeric(df1$icu)
df1 <- df1 %>% 
  mutate(race = as.numeric(race),
         race = case_when(
           race == 1 ~ 'White',
           race == 2 ~ 'Black',
           race == 3 ~ 'Asian',
           race == 4 ~ 'American Indian',
           race == 5 ~ 'Hispanic',
           race == 6 ~ 'Other',
           race == 9 ~ 'Unknown',
           
         ))
df1 <- df1 %>% 
  mutate(admtype = as.numeric(admtype),
         admtype = case_when(
           admtype == 1 ~ 'Emergency',
           admtype == 2 ~ 'Urgent',
           admtype == 3 ~ 'Electric',
           admtype == 4 ~ 'Newborn',
           admtype == 5 ~ 'Court Committal',
           admtype == 9 ~ 'Information Not Availble',
         
           ))
```


What are variables that have missing values?

```{r}
na_count <- colSums(is.na(df1))
print(na_count)
```



Remove all variables with missing values.
```{r}
df2 = drop_na(df1)
```


Refer to the data description in the file HDD2015-18cy6-20-19.docx, which variable recording the month of admission?, which variable recording the month of discharge?
```{r}
#Month of Admission = moa
#Month of Discharge = mod
```


Which month admitted the most number of patients? Which month admitted the most number of male patients?
```{r}
df1 %>%
  group_by(moa) %>%
  count() %>%
  arrange(desc(n))
```

```{r}
df1 %>%
  filter(sex == 'Male') %>%
  group_by(moa) %>%
  count(moa) %>%
  arrange(-n)
```


Which month has the most number of teenage female patients?
```{r}
df1 %>%
  filter(sex == "Female", age >= 13, age <=19 ) %>%
  group_by(moa) %>%
  count(moa) %>%
  arrange(-n)
```


Which provider has the most number of female patients in October?
```{r}
df1 %>%
  group_by(provider) %>%
  filter(sex == 'Female', moa == '10')%>%
  count(provider) %>% 
  arrange(-n)
```


Are female patients older than male patients, on average?

```{r}
df1 %>% 
  filter(sex == 'Female') %>% 
  summarise(female_age = mean(age))
```
```{r}
df1 %>% 
  filter(sex == 'Male') %>% 
  summarise(male_age = mean(age))
```

#No, female patients are not older than male patients on average.
Calculate the average age of patients by months. Which month has the oldest patients on average age?
```{r}
df1 %>% 
  group_by(moa) %>%
  summarise(average_age = mean(age)) %>% 
  arrange(-average_age)
```


What is the name of the provider that has the highest total charge?
```{r}
df1 %>% 
  group_by(provider) %>% 
  summarise(total_charge = max(total)) %>% 
  arrange(desc(total_charge)) %>% 
  head(1)
```


What is the name of the provider that has the least total charge for teenage male on average?
```{r}
df1 %>% 
  group_by(provider) %>%
  filter(sex == 'Male', age >= 13, age <=19 ) %>% 
  summarise(least_total = min(total)) %>% 
  count(provider) %>% 
  arrange(-n)
```


Create a season (Spring, Summer, Fall, Winter) variable. Calculate the length of stays by season. Which season has the longest length of stays on average?
```{r}
df1 <- df1 %>%
  mutate(season = case_when(
    month(moa) %in% c(12, 1, 2) ~ "Winter",
    month(moa) %in% c(3, 4, 5) ~ "Spring",
    month(moa) %in% c(6, 7, 8) ~ "Summer",
    month(moa) %in% c(9, 10, 11) ~ "Fall"
  ))
```

```{r}
season_average <- df1 %>%
  group_by(season) %>%
  summarise(length_of_stay = mean(los)) %>% 
  arrange(desc(length_of_stay))
```

```{r}
season_average %>%
  filter(length_of_stay == max(length_of_stay))
```

On average, how much a 20 year-old male get charged for staying 1 day in the Fall season?

```{r}
df1 %>% 
  group_by(sex == 'Male') %>% 
  filter(age == 21, season == 'Fall', los == 1) %>% 
  summarise(total = mean(total))
```

Write a paragraph to summarize the section and give your comments on the results. You could do some other calculations to support your points.

#This data provides valuable insights into hospital attendance, including the average number of people who visit, broken down by gender, age, and race. It gives a clear picture of the demographic patterns of hospital-goers.

#II. Data Visualization

Continue with the data from part I.

Provides at least 10 meaningful plots. Comments on the plots. All plots should have title, caption, appropriate labels on x and y-axis
#1 
```{r}
df1 %>% 
  ggplot() +
  geom_point(aes(x = age, y = icu, color = sex)) +
  labs(title = "Age vs. ICU Days", x = "Age", y = "ICU (Days)")
```
#2
```{r}
df1 %>% 
  ggplot() +
  geom_bar(mapping=aes(x = race, fill = sex),
  position = 'fill')+
  labs(title = "Seasonal Admissions: Based on Race", x = "Season", y = "Number of Admissions")
```

#3
```{r}
df1 %>% 
  ggplot()+
  geom_smooth(aes(x = los, y = total))+
   labs(title = "Length of Stay vs. Total Charges", x = "Length of Stay", y = "Total Charges")
```

#4
```{r}
df1 %>% 
  ggplot() +
  geom_point(aes(x = age, y = los, color = sex)) +
  labs(title = "Age vs. Length of Stay", x = "Age", y = "Length of Stay (Days)")
```

#5
```{r}
df1 %>% 
  ggplot() +
  geom_bar(mapping=aes(x = sex, fill = season),
  position = 'dodge')+
  labs(title = "Seasonal Admissions", x = "Sex", y = "Number of Admissions")
```

#6
```{r}
df1 %>% 
  ggplot() +
  geom_point(aes(x = age, y = total, color = sex)) +
  labs(title = "Age vs. Total Charges", x = "Age", y = "Total Charges")
```

#7
```{r}
ggplot(df1, aes(x = provider)) +
  geom_bar(fill = "purple", color = "blue") +
  labs(title = "Provider Frequency", x = "Provider", y = "Frequency")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

#8
```{r}
df1 %>% 
  ggplot() +
  geom_bar(mapping=aes(x = admtype, fill = sex),
  position = 'dodge')+
  labs(title = "Admission Type per Sex", x = "Admission Type", y = "Number of Admissions")+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

#9
```{r}
df1 %>%
  ggplot(aes(x = age)) +
  geom_histogram(binwidth = 5) +
  labs(title = "Age Distribution by Race", x = "Age", y = "Count") +
  facet_wrap(~race)
```

#10
```{r}
df1 %>% 
ggplot(aes(x = season, y = los)) +
  geom_boxplot(fill = "purple") +
  labs(title = "Length of Stay by Season", x = "Season", y = "Length of Stay (Days)")
```

```{r}
library(gganimate)
g <- ggplot(df1, 
       aes(x = los, 
           y = race, 
           fill = age))+
  geom_col() + 
  transition_states(sex)+
labs(title = 'Sex: {closest_state}')
animate(g)
```

## Part III

```{r}
#library(haven)
#df1 <-read_sas('hdd0318cy.sas7bdat')

total_charge <- median(df1$tot, na.rm = TRUE)

df1 <- df1 %>%
  mutate(target = case_when(
    tot < total_charge ~ "low",
    TRUE ~ "high"
  ))
df1 <- df1 %>% 
  select("age","sex","raceethn","provider","moa","mod","admtype","campus", 'los', 'target')
```

```{r}
df1 <- df1 %>%
  filter(!raceethn =='', !admtype =='')  
```

```{r}
df1 <- df1 %>%
  mutate(
    target = as.factor(target),
    raceethn = as.factor(raceethn),
    provider = as.factor(provider),
    moa = as.factor(moa),
    mod = as.factor(mod),
    admtype = as.factor(admtype),
    campus = as.factor(campus),
    age = as.numeric(age),
    los = as.numeric(los)
  )
```

```{r}
library(caret)

df1 <- drop_na(df1)

set.seed(2020)
splitIndex <- createDataPartition(df1$target, p = 0.10, list = FALSE)
df1_train <- df1[splitIndex, ]
df1_test <- df1[-splitIndex, ]
```
Question 2 

```{r}
library(rpart) 
tree_model <- rpart(target ~ ., data = df1_train,
                 control = rpart.control(maxdepth = 3))
```

```{r}
library(rattle)
fancyRpartPlot(tree_model)
```

```{r}
tree_model$variable.importance
```

```{r}
barplot(tree_model$variable.importance)
```
Question 3

```{r}
tuneGrid = expand.grid(maxdepth = 2:10)
trControl = trainControl(method = "cv",
                         number = 5)

tree_approach2 <- train(target~., data=df1_train, 
                                method = "rpart2", 
                                trControl = trControl,
                                tuneGrid = tuneGrid)

plot(tree_approach2)
```

```{r}
library(caret)
pred <- predict(tree_approach2, df1_test)

cm <- confusionMatrix(data = pred, reference = df1_test$target)
cm$overall[1]
```

```{r}
library(randomForest)
forest_model = randomForest(target ~ ., data=df1_train, ntree = 5)

pred <- predict(forest_model, df1_test, type = "class")

cm <- confusionMatrix(data = pred, reference = df1_test$target, positive = "high")
cm$overall[1]
```
Question 4

```{r}
trControl = trainControl(method = "cv",
                         number = 5)
tree <- train(target~., data=df1_train, 
                                method = "rpart2", 
                                trControl = trControl)
forest_ranger <- train(target~., data=df1_train, 
                    method = "ranger", 
                                trControl = trControl)

results <- resamples(list('Decision Tree' = tree,
                          'Random Forest' = forest_ranger))
bwplot(results)
```
Question 5

```{r}
pred <- predict(tree, df1_test)
cm <- confusionMatrix(data = pred, reference = df1_test$target, positive = "high")
cm$overall[1]
```
Question 6

```{r}
median(df$age)
```


```{r}
library(haven)
library(gganimate)
library(ggplot2)
library(tidyverse)
df2 = df
```

```{r}
df_clean <- na.omit(df2)
```

```{r}
df_clean <- df[complete.cases(df$age), ]
```

```{r}
median(df$age)
```


```{r}
df2$target <- ifelse(df2$age<56,'low','high')
df2 <- df2 %>%
  mutate(target=as.factor(target))
df2 = df2 %>% select("age","sex","provider","moa","mod","campus","los","target")
```


#1
```{r}
library(caret)

splitIndex <- createDataPartition(df2$target, p = 0.10, list = FALSE)
df1_train <- df2[splitIndex, ]
df1_test <- df2[-splitIndex, ]
```

#2
```{r}
library(rpart) 
tree_model <- rpart(target ~ ., data = df1_train,
                 control = rpart.control(maxdepth = 3))
```

```{r}
library(rattle)
fancyRpartPlot(tree_model)
```

```{r}
tree_model$variable.importance
```
```{r}
barplot(tree_model$variable.importance)
```
#3
```{r}
library(caret)

sum(is.na(df1_train))  
sum(is.na(df1_train$target))  

df1_train_clean <- na.omit(df1_train)

df1_train_clean$target <- as.factor(df1_train_clean$target)

tuneGrid = expand.grid(maxdepth = 2:10)
trControl = trainControl(method = "cv", number = 5)

tree_approach2 <- train(target ~ ., data = df1_train_clean,
                        method = "rpart2", 
                        trControl = trControl,
                        tuneGrid = tuneGrid)

plot(tree_approach2)

```






